---
title: "Lab5"
author: "Lewis White"
date: "2023-02-07"
output: html_document
---

This week's lab is a musical lab. You'll be requesting data from the Spotify API and using it to build k-nearest neighbor and decision tree models. In order to use the Spotify you must have a Spotify account. If you don't have one, sign up for a free one here: <https://www.spotify.com/us/signup>.

Once you have an account, go to Spotify for developers (<https://developer.spotify.com/>) and log in. Click the green "Create a Client ID" button to fill out the form to create an app create an app so you can access the API. On your developer dashboard page, click on the new app you just created. On the app's dashboard page you will find your Client ID just under the header name of your app. Click "Show Client Secret" to access your secondary Client ID. When you do this you'll be issued a Spotify client ID and client secret key.

You have two options for completing this lab.

**Option 1**: **Classify by users**. Build models that predict whether a given song will be in your collection vs. a partner in class. This requires that you were already a Spotify user so you have enough data to work with. You will download your data from the Spotify API and then exchange with another member of class.

-   **I am selecting this option.**

**Option 2**: **Classify by genres**. Build models that predict which genre a song belongs to. This will use a pre-existing Spotify dataset available from Kaggle.com (<https://www.kaggle.com/datasets/mrmorj/dataset-of-songs-in-spotify>)

```{r}
#LOAD NECESSARY PACKAGES

library(spotifyr) #API interaction
library(tidyverse) #cleaning data, ggplot, etc 
library(tidymodels) #for modeling/statistical analysis
library(rsample) #for splitting data into train / test
library(recipes) #for creating the recipe for ML
#library(skimr) #for data exploration / early summary stats and viz
library(kknn) #for KNN modeling
library(plotly) #for data viz
library(ggpubr) #for data viz
library(here) #for simplifying file path navigation
library(baguette) #for bagging decision trees
library(ranger) # engine for random forests
library(kableExtra) #for creating a formatted table
```

Client ID and Client Secret are required to create and access token that is required to interact with the API. You can set them as system values so we don't have to do provide them each time.

```{r, eval = FALSE}
#API key stored in token.R, which is not git tracked for security

source(here("token.R"))
```

This may result in an error: "INVALID_CLIENT: Invalid redirect URI." This can be resolved by editing the callback settings on your app. Go to your app and click "Edit Settings". Under redirect URLs paste this: <http://localhost:1410/> and click save at the bottom.

### **Data Preparation**

You can use get_my_saved_tracks() to request all your liked tracks. It would be good if you had at least 150-200 liked tracks so the model has enough data to work with. If you don't have enough liked tracks, you can instead use get_my_recently_played(), and in that case grab at least 500 recently played tracks if you can.

The Spotify API returns a dataframe of tracks and associated attributes. However, it will only return up to 50 (or 20) tracks at a time, so you will have to make multiple requests. Use a function to combine all your requests in one call.

```{r}
#GETTING A DATAFRAME OF LIKED SONGS

#In the get_my_saved_tracks() function, you can use offsets() to specify the index of the first track to return. 
offsets = seq(from = 0, to = 1000, by = 50)

#initializing an empty matrix 
liked_tracks <- data.frame(matrix(nrow = 0, ncol = 30))

#Function to get my 1050 most recently liked tracks 
for (i in seq_along(offsets)) {  
  my_tracks = get_my_saved_tracks(limit = 50, 
                                     offset = offsets[i])
  df_temp = as.data.frame(my_tracks) #creating a temporary data frame to store the 50 liked tracks from a given run
  liked_tracks <- rbind(liked_tracks, df_temp) #binding the temporary data frame to my liked tracks data frame. 
}
```

Once you have your tracks, familiarize yourself with this initial dataframe. You'll need to request some additional information for the analysis. If you give the API a list of track IDs using get_track_audio_features(), it will return an audio features dataframe of all the tracks and some attributes of them.

```{r}
#obtain a list of the track IDs 
ids <- liked_tracks$track.id

#the ids argument in `get_track_audio_features()` can only take 100 IDs at a time, so I'm splitting them into 100 track groupings
ids_split <- split(ids, ceiling(seq_along(ids) / 100))

#create an empty list
my_tracks_audio_feats_list <- list()

#Iterating the `get_track_audio_features()`function over each ID split and storing the resulting data in a list.
for (i in 1:length(ids_split)) {
  my_tracks_audio_feats_list[[i]] <- get_track_audio_features(ids = ids_split[[i]])
}

#Combine the list of data frames into a single data frame.
my_tracks_audio_feats <- do.call(rbind, my_tracks_audio_feats_list)
```

These track audio features are the predictors we are interested in, but this dataframe doesn't have the actual names of the tracks. Append the 'track.name' column from your favorite tracks database.

```{r}
#selecting the track id and track name from liked tracks so I can use left_join and only add the track name to the audio features data set
liked_tracks_join <- liked_tracks %>%
  select(track.id, track.name, track.artists) %>%
  mutate(primary_artist = unlist(lapply(liked_tracks$track.artists, function(x) x$name[1]))) %>%
  select(-track.artists)

#add the track name and artist to the audio features data frame 
liked_tracks_full <- left_join(my_tracks_audio_feats, liked_tracks_join, by = c("id" = "track.id"))

#save this data frame as a CVS so it can be shared with others
write_csv(liked_tracks_full, "lewis_liked_tracks.csv")
```

Find a class mate whose data you would like to use. Add your partner's data to your dataset. Create a new column that will contain the outcome variable that you will try to predict. This variable should contain two values that represent if the track came from your data set or your partner's.

```{r}
# reading in Elke's liked tracks. Add a column to make it clear which tracks are Elke's  
elke_liked_tracks <- read_csv(here("elke_liked_tracks.csv")) %>%
  mutate(listener = "Elke")

#add a column named "listener" to make it clear that these are my tracks 
lewis_liked_tracks <- liked_tracks_full %>%
  mutate(listener = "Lewis")

# Combine all of our liked tracks into one data frame 
all_tracks <- rbind(lewis_liked_tracks, elke_liked_tracks) %>%
  select(-(type:analysis_url))
```

### Data Exploration

Let's take a look at your data. Do some exploratory summary stats and visualization.

For example: What are the most danceable tracks in your dataset? What are some differences in the data between users (Option 1) or genres (Option 2)?

#### Lists of songs 

```{r}
# names(lewis_liked_tracks)  #for checking out column names

#my most danceable songs 
lewis_liked_tracks %>%
  select(danceability, track.name, primary_artist) %>%
  arrange(-danceability) %>%
  head(5)
  
#my least danceable songs
lewis_liked_tracks %>%
  select(danceability, track.name, primary_artist) %>%
  arrange(danceability) %>%
  head(5)

#my most "speechy" songs
lewis_liked_tracks %>%
  select(speechiness, track.name, primary_artist) %>%
  arrange(-speechiness) %>%
  head(5)
  
#my least "speechy" songs
lewis_liked_tracks %>%
  select(speechiness, track.name, primary_artist) %>%
  arrange(speechiness) %>%
  head(5)
```

#### Distribution Plots

```{r}
#Danceability comparison
dance_plot <- ggplot(all_tracks, aes(x = danceability, fill = listener,
                    text = paste(listener))) +
  geom_density(alpha=0.6, color=NA) +
  scale_fill_manual(values=c("#b0484f", "#4436d9"))+
  labs(x="Danceability", y="Density") +
  guides(fill=guide_legend(title="Listener"))+
  theme_minimal() +
  ggtitle("Distribution of Danceability Data")


#speechiness comparison
speech_plot <- ggplot(all_tracks, aes(x = speechiness, fill = listener,
                    text = paste(listener))) +
  geom_density(alpha=0.6, color=NA) +
  scale_fill_manual(values=c("#b0484f", "#4436d9"))+
  labs(x="Speechiness", y="Density") +
  guides(fill=guide_legend(title="Listener"))+
  theme_minimal() +
  ggtitle("Distribution of Speechiness Data")


#acousticness comparison
acoustic_plot <- ggplot(all_tracks, aes(x = acousticness, fill = listener,
                    text = paste(listener))) +
  geom_density(alpha=0.6, color=NA) +
  scale_fill_manual(values=c("#b0484f", "#4436d9"))+
  labs(x="Acousticness", y="Density") +
  guides(fill=guide_legend(title="Listener"))+
  theme_minimal() +
  ggtitle("Distribution of Acousticness Data")

#energy comparison 
energy_plot <- ggplot(all_tracks, aes(x = energy, fill = listener,
                    text = paste(listener))) +
  geom_density(alpha=0.6, color=NA) +
  scale_fill_manual(values=c("#b0484f", "#4436d9"))+
  labs(x="Energy", y="Density") +
  guides(fill=guide_legend(title="Listener"))+
  theme_minimal() +
  ggtitle("Distribution of Energy Data")

ggarrange(dance_plot, speech_plot, acoustic_plot, energy_plot, ncol=2, nrow=2, common.legend = TRUE, legend="bottom")

#My music appears to be higher energy, more danceable, and more "speechy," while Elke's music tends to be more acoustic. 
```

```{r, warning = FALSE}
#plotting valence and energy to get a sense for the moods of our liked songs 
ggplot(data = all_tracks, aes(x = valence, y = energy, color = listener)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 1)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  annotate('text', 0.25 / 2, 0.95, label = "Turbulent/Angry", fontface =
             "bold") +
  annotate('text', 1.75 / 2, 0.95, label = "Happy/Joyful", fontface = "bold") +
  annotate('text', 1.75 / 2, 0.05, label = "Chill/Peaceful", fontface =
             "bold") +
  annotate('text', 0.25 / 2, 0.05, label = "Sad/Depressing", fontface =
             "bold") +
  theme_minimal() +
  labs(x = "Valence", 
       y = "Energy",
       title = "Plotting songs based on their positivity and energy level",
       subtitle = "Elke and I don't have many songs in the Chill/Peaceful quadrant.")


## this plotly allows you to hover over points to see the track and the corresponding artist
plot_ly(data = all_tracks,
        x = ~valence, y = ~energy, color = ~listener, type = "scatter", mode = "markers",
        text = paste("Track Name:", all_tracks$track.name, "<br>",
                     "Primary Artist:", all_tracks$primary_artist, "<br>",
                     "Valence:", all_tracks$valence, "<br>", 
                     "Energy:", all_tracks$energy, "<br>",
                     "Listener:", all_tracks$listener)) %>%
  layout(xaxis = list(title = "Valence"),
         yaxis = list(title = "Energy"),
         hovermode = "closest",
         title = "Track Valence vs Energy")
```

#### Statistical comparison of variables 

```{r}
#some quick t-tests comparing my music to Elke's 

#danceability 
t.test(lewis_liked_tracks$danceability, elke_liked_tracks$danceability, var.equal = FALSE)

#speechiness
t.test(lewis_liked_tracks$speechiness, elke_liked_tracks$speechiness, var.equal = FALSE)

#acousticness
t.test(lewis_liked_tracks$acousticness, elke_liked_tracks$acousticness, var.equal = FALSE)

#energy
t.test(lewis_liked_tracks$energy, elke_liked_tracks$energy, var.equal = FALSE)

#Each t.test was significant at the 0.001 significance level, clearly suggesting that my music is different than Elke's across the variables tested. 
```

### **Modeling**

Create two models, a k-nearest neighbor model and a decision tree model that predict whether a track belongs to you or your partner's collection. Then validate and compare the performance of the two models you have created.

Make sure to use appropriate resampling to select the best version of each algorithm to compare and some appropriate visualization of your results.

```{r}
#prepare all_tracks for modeling by making nominal data a factor and removing variables that don't make sense to include in the model
all_tracks_modeling <- all_tracks %>% 
  mutate_if(is.ordered, .funs = factor, ordered = F) %>%
  select(-track.name, -primary_artist) %>%
  mutate(listener = as.factor(listener))
```

```{r}
#splitting the data
set.seed(123)
#initial split of data ~ we're going with 70/30 because the sample size isn't super large for testing
tracks_split <- initial_split(all_tracks_modeling, prop = .7)
tracks_train <- training(tracks_split)
tracks_test <- testing(tracks_split)

#pre-processing
# We need to create a recipe and do the pre-processing by converting dummy coding the nominal variables and normalizing the numeric variables.
tracks_recipe <- recipe(listener ~ ., data = tracks_train) %>%
  #step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_normalize(all_numeric(), -all_outcomes()) %>% #normalize numeric to make sure scale is okay
  prep()
```

### KNN Model

```{r spec_with_tuning}
# Define our KNN model with tuning
knn_spec_tune <- nearest_neighbor(neighbor = tune()) %>%
set_mode("classification") %>%
set_engine("kknn")

# Check the model
knn_spec_tune
```

```{r}
# Define a new workflow
wf_knn_tune <- workflow() %>%
add_model(knn_spec_tune) %>%
add_recipe(tracks_recipe)

# 10-fold CV on the training dataset
set.seed(123)
cv_folds <- tracks_train %>%
  vfold_cv(v = 10)


# Fit the workflow on our predefined folds and hyperparameters
fit_knn_cv <- wf_knn_tune %>%
  tune_grid(
    cv_folds, 
    grid = data.frame(neighbors = c(1, 5, seq(10, 150, 10))) # try with 1 nearest neighbor, try with 5, 10, 20, 30, ..., 100
  )
    
# Check the performance with collect_metrics()
print(n=32, fit_knn_cv %>% collect_metrics()) 
```

```{r}
# The final workflow for our KNN model
final_wf <-
  wf_knn_tune %>%
  finalize_workflow(select_best(fit_knn_cv))

# Check out the final workflow object
final_wf

# Fitting our final workflow
final_fit <- final_wf %>%
  fit(data = tracks_train)

# Examine the final workflow
final_fit
```

```{r}
#generating predictions using the model on the test data
tracks_pred <- final_fit %>% predict(new_data = tracks_test)

tracks_pred %>% head()

# Write over 'final_fit' with this last_fit() approach
final_fit <- final_wf %>% last_fit(tracks_split)

# Collect metrics on the test data!
final_fit %>% collect_metrics()
```

### Decision Tree Approach

```{r}
#new spec, tell the model that we are tuning hyperparams
tree_spec_tune <- decision_tree(
  cost_complexity = tune(),   #tune() asks R to try a bunch of different parameters. 
  tree_depth = tune(),
  min_n = tune()) %>%
  set_engine("rpart") %>%
  set_mode("classification")


tree_grid <- grid_regular(cost_complexity(), tree_depth(), min_n(), levels = 5) #grid_regular shows grid of different input tuning options
#levels says how many combinations we should try. 

tree_grid  #125 options 
```

```{r}
wf_tree_tune <- workflow() %>%
  add_recipe(tracks_recipe) %>%
  add_model(tree_spec_tune)
    #when you add recipe into workflow, it automatically will prep and bake when necessary. 
  

#workflow pulls together the specification. and then we can fit on the workflow. you could fit the wf_tree_tune 
```

```{r}
#set up k-fold cv. This can be used for all the algorithms. I switch to 5 fold CV instead of 10 (used in KNN above) for computation speed. 
listener_cv <- tracks_train %>% vfold_cv(v=5) #creating 5 folds cross validation 
listener_cv
```

```{r}
doParallel::registerDoParallel() #build trees in parallel
#200s

tree_rs <- tune_grid(
  wf_tree_tune,
  listener ~ ., #function 
  resamples = listener_cv, #resamples to use
  grid = tree_grid, #grid to try
  metrics = metric_set(accuracy) #how to assess which combinations are best 
)

tree_rs
```

```{r}
autoplot(tree_rs) + theme_light()
```

```{r select_hyperparam}
show_best(tree_rs)
select_best(tree_rs)
```

```{r final_tree_spec}
final_tree <- finalize_model(tree_spec_tune, select_best(tree_rs))
```

This model has not been fit yet though.

```{r final_tree_fit}
final_tree_fit <- last_fit(final_tree, listener ~., tracks_split)

final_tree_fit$.predictions
```

### Bagging Trees

```{r}
set.seed(123)
bag_spec_tune <- bag_tree(cost_complexity = tune(),
                          tree_depth = tune(),
                          min_n = tune()) %>%
  set_mode("classification") %>%
  set_engine("rpart", times = 50)

bag_grid <- grid_regular(cost_complexity(), tree_depth(), min_n(), levels = 5)

bag_grid


wf_bag_tune <- workflow() %>% 
  add_recipe(tracks_recipe) %>%
  add_model(bag_spec_tune)

doParallel::registerDoParallel() #build trees in parallel
#200s

bag_rs <- tune_grid(
  bag_spec_tune,
  listener ~ ., #function 
  resamples = listener_cv, #resamples to use
  grid = bag_grid,
  metrics = metric_set(accuracy) #how to assess which combinations are best 
)

bag_rs

autoplot(bag_rs) + theme_light()

show_best(bag_rs)
select_best(bag_rs)
```

```{r final_tree_spec}
final_bag <- finalize_model(bag_spec_tune, select_best(bag_rs))
```

This model has not been fit yet though.

```{r final_tree_fit}
final_bag_fit <- last_fit(final_bag, listener ~., tracks_split)

final_bag_fit$.predictions
```

### Random Forests

4.  random forest
    -   rand_forest()
    -   m_try() is the new hyperparameter of interest for this type of model. Make sure to include it in your tuning process

Go through the modeling process for each model: \* Preprocessing. You can use the same recipe for all the models you create. \* Resampling. Make sure to use appropriate resampling to select the best version created by each algorithm. \* Tuning. Find the best values for each hyperparameter (within a reasonable range).

Compare the performance of the four final models you have created.

Use appropriate performance evaluation metric(s) for this classification task. A table would be a good way to display your comparison. Use at least one visualization illustrating your model results.

```{r}
set.seed(123)
# Bagging specifications
forest_spec_tune <- 
  rand_forest(min_n = tune(),
  mtry = tune(),
  trees =  140) %>% # 14 predictors * 10
  set_engine("ranger") %>%
  set_mode("classification")

forest_grid <- grid_regular(mtry(c(1,13)), min_n(), levels = 5)

forest_grid

wf_forest_tune <- workflow() %>% 
  add_recipe(tracks_recipe) %>%
  add_model(forest_spec_tune)

doParallel::registerDoParallel() #build forest in parallel

forest_rs <- tune_grid(
  forest_spec_tune,
  listener ~ ., 
  resamples = listener_cv, #resamples to use
  grid = forest_grid,
  metrics = metric_set(accuracy) #how to assess which combinations are best 
)

forest_rs

autoplot(forest_rs) + theme_light()

show_best(bag_rs)
select_best(bag_rs)
```

```{r final_tree_spec}
final_forest <- finalize_model(forest_spec_tune, select_best(forest_rs))

final_forest_fit <- last_fit(final_forest, listener ~., tracks_split)

final_forest_fit$.predictions
final_forest_fit$.metrics
```

### Comparing each of the models

```{r}
#cleaning model metrics to create a table with each model's performance metrics 

forest_results_table <- final_forest_fit %>% collect_metrics() %>%
  select(.metric, .estimate) %>%
  mutate(method = "Random Forest")

bagging_results_table <- final_bag_fit %>% collect_metrics() %>%
  select(.metric, .estimate) %>%
  mutate(method = "Bagging")

decision_tree_results_table <- final_tree_fit %>% collect_metrics() %>%
  select(.metric, .estimate) %>%
  mutate(method = "Decision Tree")

knn_results_table <- final_fit %>% collect_metrics() %>%
  select(.metric, .estimate) %>%
  mutate(method = "KNN")

majority_class_results_table <- data.frame (.metric  = c("accuracy", "roc_auc"),
                  .estimate = c(nrow(lewis_liked_tracks) / (nrow(lewis_liked_tracks) + nrow(elke_liked_tracks)), .5),
                  method = c("dummy", "dummy")
                  )

full_results <- bind_rows(majority_class_results_table,
                          knn_results_table, 
                          decision_tree_results_table,
                          bagging_results_table,
                          forest_results_table) %>%
  select(method, .estimate, .metric)

#table of results 
full_results %>%
  kbl(caption = "Table 1. Classification Methods and their Respective Accuracies") %>%
  kable_classic(full_width = T, html_font = "Cambria")
  
```

```{r}
#accuracy plot

ggplot(data = full_results, aes(x = method, y = .estimate, fill = .metric)) +
  geom_bar(stat='identity', position='dodge') + 
  theme_minimal() +
  scale_fill_manual(values=c("#9fd182", "#7798c9")) +
  labs(y = "Accuracy Estimate",
       x = "Classification Method",
       fill = "Accuracy Metric", 
       title = "Random Forest Classification Performed the Best Across Both Accuracy Metrics")
```

### 
